Competing Bandits: The Perils of Exploration under Competition

### 2018-09-04 paper/content/inverted_u.tex
fixups to paper
use more game theoretic terminology

Now we demonstrate the inverted-U relationship between the
incentivized algorithms and the competition levels. In particular, we
investigate the incentivized algorithms under three different regimes:
permanent monopoly, temporary monopoly, and permanent duopoly: \swcomment{think the tie-breaking should be discussed in the prelim?}

\begin{finding}
\textit{For sufficiently low warm start under the instances that are relative reputation costly, the optimal strategies in the competition game are:
\begin{center}
\textbf{Permanent Monopoly} - $\DG$ \\
\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
\textbf{Permanent Duopoly} - $\DG$ \\
\end{center}
The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
}
\end{finding}

\xhdr{Permanent Monopoly.} Since there is only a single firm in the
market for the entire period, the firm can take the entire market
regardless of what algorithm it deploys. \gadelete{Assuming the
  deployment of any exploration algorithms will incur certain cost}
\swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$)
  would incur deployment cost}, the firm would choose to deploy $\DG$.

@@

\xhdr{Permanent Duopoly} \swdelete{Permanent duopoly corresponds to
  the model described in section 2 where $X = 0$ so that both firms
  enter the market simultaneously. What algorithms are firms
  incentivized to deploy? We report the average market share taken
  over the $N$ simulations for a set of exogenous values of
  $T_0$. Additionally, we report the mean and median of the effective
  end of game.}  \swedit{Since both firms enter at the same time, they
  will both get an expected market share of $1/2$ whenever they deploy
  the same algorithms. We will mainly investigate the market shares
  when the firms play different algorithms.  In Tables \ref{sim_ht}
  and \ref{sim_nih}, we report the average market share taken over the
  $N$ simulations for different values of warm-up length $T_0$, and
  also the mean and median of the EEOG, under heavy tail and needle in
  haystack instances.}

For the relative reputation costly instances,\footnote{We defer the table for uniform instances to the appendix.} we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$. This establishes our claim that $\DG$ is the incentivized algorithm.

\gadelete{
\begin{puzzle}\label{puz:dg_ht}
Even for large warm start, $\DG$ is better than $\DEG$ under heavy tail instances.\footnote{We demonstrate this holds even for $K = 3$ in the supplementary material}
\end{puzzle}

\begin{puzzle}\label{puz:nih}
Under Needle In Haystack, $\TS$ is weakly dominant across all warm starts. 
\end{puzzle}
}

@@

\xhdr{Temporary Monopoly.} Recall that the incumbent firm enters the
market and serves as a monopolist until the entrant firm enters at
round $X$. \gaedit{For this section we focus on the warm start values and on the relative reputation costly instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy. Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary
  material. The results are unsurprising, where $\TS$ is the dominant
  strategy for both the entrant and incumbent.}.

>>>>

Now we demonstrate the inverted-U relationship between the
incentivized algorithms and the competition levels. In particular, we
investigate the incentivized algorithms under three different regimes:
permanent monopoly, temporary monopoly, and permanent duopoly:

\begin{finding}
\textit{For sufficiently low warm start under the instances that are relative reputation costly, the unique pure Nash equilibrium strategies in the competition game are:\footnote{the uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
\begin{center}
\textbf{Permanent Monopoly} - $\DG$ \\
\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
\textbf{Permanent Duopoly} - $\DG$ \\
\end{center}
The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
}
\end{finding}

\xhdr{Permanent Monopoly.} Since there is only a single firm in the
market for the entire period, the firm can take the entire market
regardless of what algorithm it deploys. \swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$) would incur deployment cost and firms break indifferences towards easier to deploy algorithms}, the firm would choose to deploy $\DG$.

@@

\xhdr{Permanent Duopoly} \swedit{Since both firms enter at the same time, they will both get an expected market share of $1/2$ whenever they deploy
  the same algorithms. We will mainly investigate the market shares
  when the firms play different algorithms.  In Tables \ref{sim_ht}
  and \ref{sim_nih}, we report the average market share taken over the
  $N$ simulations for different values of warm-up length $T_0$, and
  also the mean and median of the EEOG, under heavy tail and needle in
  haystack instances.}

For the relative reputation costly instances,\footnote{We defer the table for uniform instances to the appendix. Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG. Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium} we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$. This establishes our claim that $\DG$ is the incentivized algorithm.

@@

\xhdr{Temporary Monopoly.} Recall that the incumbent firm enters the
market and serves as a monopolist until the entrant firm enters at
round $X$. \gaedit{For this section we focus on the warm start values and on the relative reputation costly instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy. Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary
  material. The results are unsurprising, where $\TS$ is the dominant
  strategy for both the entrant and incumbent.}

### 2018-09-04 paper/content/model.tex
fixups to paper
use more game theoretic terminology

\item an exploration-separated algorithm that in each round, \emph{explores} with probability $\eps$: chooses an arm independently and uniformly at random, and with the remaining probability \emph{exploits} according to \DG. We call it ``dynamic epsilon-greedy", \DEG in short.

@@

\swedit{For any fixed firm, we say that algorithm 1 \emph{weakly
    dominates} algorithm 2 if algorithm 1 never provides a lower
  market share than than 2 against all opposing strategies. An
  algorithm is a weakly dominant strategy for a firm if the algorithm
  weakly dominates all other algorithms.}

\swcomment{please use the above terminology}

>>>>

\item an exploration-separated algorithm that in each round, \emph{explores} with probability $\eps$: chooses an arm independently and uniformly at random, and with the remaining probability \emph{exploits} according to \DG. We call it ``dynamic epsilon-greedy", \DEG in short.\gaedit{\footnote{For our experiments, we fix $\eps = 0.05$. Our pilot experiments showed that different $\eps$ did not qualitatively change the results.}}

@@

\swedit{For any single firm, we say that algorithm 1 \emph{weakly
    dominates} algorithm 2 if algorithm 1 never provides a lower
  market share than 2 against all opposing strategies. An
  algorithm is a weakly dominant strategy for a firm if the algorithm
  weakly dominates all other algorithms.}
  \gaedit{Since the algorithms that involve exploration have larger deployment costs, we break any indifference towards algorithms that do not involve exploration (i.e. if \DG and \DEG provide the same market share, the firm will play \DG).}
\swcomment{please use the above terminology}
